|

### 2025-04-29 manuscript/paper.Rmd
paper updated.

In our setting, we replace $X$ and $Y$ with $r(\boldsymbol{\theta})$ and $\hat{r}(T(\phi, s),\boldsymbol{\theta})$, and we could seek to estimate $I(r(\boldsymbol{\theta}),\hat{r}(T(\phi, s),\boldsymbol{\theta}))$.  We note here that $r(\boldsymbol{\theta})$ is not a random variable, however, since the indexing of the teams in the vector $\boldsymbol{\theta}$ is arbitrary (we use 1, 2, 3, etc. for convenience, but any indexing is valid), we can view this a random variable where all permutation so of the $n$ teams are equally likely.  Therefore when calculating  the mutual information in this setting, we set $p(r(\boldsymbol{\theta})) = \frac{1}{n!}$.  By a similar argument we set $\hat{r}(T(\phi, s),\boldsymbol{\theta}) = \frac{1}{n!}$.  This

@@

# ryan way 
Consider a set of $n$ teams indexed from $i = 1, 2, \cdots, n$ each with an associated true strength parameter $\boldsymbol{\theta} = (\theta_1, \theta_2, \ldots, \theta_n)$ such that $(\theta_1 < \theta_2 < \ldots < \theta_n)$.  Next, let $r(\boldsymbol{\theta})$ be the rank of the vector $\boldsymbol{\theta}$ from largest to smallest so that $r(\boldsymbol{\theta}) = \{n, n-1, \ldots, 2, 1\}$.  Next define $T(\phi, s)$ as a tournament with schedule $\phi$ and seeding structure $s$.  $\phi$ contains all the information about the scheduling of teams, which could be fully known prior to the tournament (i.e. round robin) or determined as the tournament progresses (i.e. single elimination tournament).  We then let  $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ be a vector valued random variable that gives the results of a tournament as a vector of ranks.

Figure out what $\hat{r}$ should look like.

@@

Let $\boldsymbol{x} = (x_1, x_2, \ldots, x_n) $ be the permutation of indexes of $\boldsymbol{\theta}$ such that $x_1 = {i | \theta_i = max(\boldstyle{\theta})}$

>>>>

In our setting, we replace $X$ and $Y$ with $r(\boldsymbol{\theta})$ and $\hat{r}(T(\phi, s),\boldsymbol{\theta})$, and we could seek to estimate $I(r(\boldsymbol{\theta}),\hat{r}(T(\phi, s),\boldsymbol{\theta}))$.  For simplicity, we drop the arguments from the functions $r$ and $\hat{r}$ for ease of exposition.

We note here that $r$ is not a random variable, however, since the indexing of the teams in the vector $\boldsymbol{\theta}$ is arbitrary (we use 1, 2, 3, etc. for convenience, but any indexing is valid), we can view this a random variable where all permutation so of the $n$ teams are equally likely.  Therefore when calculating  the mutual information in this setting, we set $p(r) = \frac{1}{n!}$.  By a similar argument we set $p(\hat{r}) = \frac{1}{n!}$.

We define the mutual information of $r$ and $\hat{r}$ to be:

$$
I(r,\hat{r})=\sum_{r}\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})} = n!\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{(\frac{1}{n!})^2}
$$.

While there are $n!$ different permutations for the result of $r$, we don't need to sum across these as any specific choice of $r$ is just an arbitrary labeling of the true strength parameters vector $\boldsymbol{\theta}$.  So given $r$, the distribution of $\hat{r}$ is the same up to the labeling.  Therefore, we only need to consider a single permutation of $r$, we compute the the quantity inside the summation, and then multiple by $n!$ (effectively summing across $r$).

In order to compute this quantity, we need to compute $p(r,\hat{r})$.  This probability is found by calculating the probability of a given permutation of $\hat{r}$ and $r$ is assumed to be the permutation from $\{1, 2, \ldots, n\}$.  We estimate $p(r,\hat{r})$ empirically through simulation.

However, using mutual information in this form does not suit our needs in this setting.  The problem is that mutual information in this form will yield high values of mutual information when the output from the tournament $\hat{r}$ is highly consistent *even if the ranking from the tournament is incorrect*.  As an example, if $n=4$ and the true order of \boldsymbol{\theta} is $r = \{1, 2, 3, 4\}$ and $p(\hat{r} = \{4, 3, 2, 1\}) = 1$ will be the same mutual information as when $r = \{1, 2, 3, 4\}$ and $p(\hat{r} = \{1, 2, 3, 4\}) = 1$ and in both of these cases the mutual information will be maximized at:

$$
I(r,\hat{r})= n!\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{\left(\frac{1}{n!}\right)^2} = n! log((n!)^2)
$$
and for the specific case when $n=4$ would be: 
$$
4! log_2(4!^2) = 24*log_2(24^2) = 220.0782
$$.

In order to alleviate this problem, we instead consider *weighted* mutual information.  We want to give more weight to permutations from $\hat{r}$ that are "closer" to $r$.  There are any number of reasonable choices for this weighting function, and we choose to use squared error.  Specifically, 
$$
w(r, \hat{r}) =

\left\{  
\begin{array}{cc}
    \begin{array}{cc}
      \frac{1}{\hat{r}'r} & r \ne \hat{r} \\
      1 & r = \hat{r} \\
    \end{array}
\end{array}
\right. 
$$
In general, though any loss function $l(r, \hat{r})$ can be used:
$$
w(r, \hat{r}) =

\left\{  
\begin{array}{cc}
    \begin{array}{cc}
      \frac{1}{l(r, \hat{r})} & r \ne \hat{r} \\
      1 & r = \hat{r} \\
    \end{array}
\end{array}
\right. 
$$

@@

<!-- # ryan way  -->
<!-- Consider a set of $n$ teams indexed from $i = 1, 2, \cdots, n$ each with an associated true strength parameter $\boldsymbol{\theta} = (\theta_1, \theta_2, \ldots, \theta_n)$ such that $(\theta_1 < \theta_2 < \ldots < \theta_n)$.  Next, let $r(\boldsymbol{\theta})$ be the rank of the vector $\boldsymbol{\theta}$ from largest to smallest so that $r(\boldsymbol{\theta}) = \{n, n-1, \ldots, 2, 1\}$.  Next define $T(\phi, s)$ as a tournament with schedule $\phi$ and seeding structure $s$.  $\phi$ contains all the information about the scheduling of teams, which could be fully known prior to the tournament (i.e. round robin) or determined as the tournament progresses (i.e. single elimination tournament).  We then let  $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ be a vector valued random variable that gives the results of a tournament as a vector of ranks.   -->

<!-- Figure out what $\hat{r}$ should look like.   -->

@@

Let $\boldsymbol{x} = (x_1, x_2, \ldots, x_n)$ be the permutation of indexes of $\boldsymbol{\theta}$ such that $x_1 = {i | \theta_i = max(\boldsymbol{\theta})}$
